Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-08 mr-review.Rmd
added text about correlated horizontal pleiotropy

Recognition of complex traits' diverse genetic architectures has fueled development of two-sample MR 
methods with multiple SNP instruments [@hormozdiari_leveraging_2018;@watanabe_global_2019].

@@

In the case of continuous exposures and outcomes, we extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, we perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second regression involves regressing the outcome variable $Y$ on 
the fitted values.

@@

@@

@yuan_testing_2020

>>>>

Recognition of complex traits' diverse genetic architectures has fueled development of two-sample MR 
methods with multiple SNP instruments [@hormozdiari_leveraging_2018;@watanabe_global_2019;@pierce_power_2011;@bowden_mendelian_2015].

@@

In the case of continuous exposures and outcomes, we extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, we perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second stage involves regressing the outcome variable $Y$ on 
the fitted values, $\hat X$.

@@

Researchers have recognized two types of horizontal pleiotropy [@yuan_likelihood_2021]. The first occurs via 
exposure-independent paths. The resulting horizontal pleiotropic effects are independent of the
SNP-exposure relationships. The second type of horizontal pleiotropy manifests when 
in the presence of unobserved exposure-outcome confounding. It induces correlation between 
horizontal pleiotropic effects and SNP-exposure effects. Both types of horizontal pleiotropy 
violate standard MR modeling assumptions and can bias causal effect estimates and can increase false 
discoveries [@yuan_likelihood_2021].

Early MR methods account manage horizontal pleiotropy by removing instrumental SNPs 
that may be associated with the outcome variable [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].

@@

The second type of horizontal pleiotropy, from above, is sometimes called "correlated horizontal pleiotropy". Correlated horizontal pleiotropy occurs when 
CAUSE and MRAID are two methods that aim to model and to account for correlated horizontal pleiotropy.
